\Sof{TBA - New Titles in the comments}

### 2020-05-07 paper/icsme20.tex
conclusion update

This work presents an empirical study on the impact of $998$ security
patches on the maintainability of $262$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in previous work~\cite{Olivari:2018, cruz2019energyoriented}. Security patches 
significantly affect the software maintainability of open-source software, i.e., 
developers hinder maintainability when patching vulnerabilities. We found this evidence in
$41.28\%$ of the studied cases. Regarding regular commits, it seems that overall
maintainability stays the same ($42.69\%$). However, there is still a 
considerable amount of cases that harm maintainability ($27.05\%$).

This study also shows evidence that there are security patterns that need more
attention than others. \emph{Improper Input Validation} (CWE-20), \emph{Information 
Exposure} (CWE-200), \emph{Improper Restriction of Operations within the Bounds of a 
Memory Buffer} (CWE-119), \emph{Missing Release of Memory after Effective Lifetime} 
(CWE-401) and \emph{Path Traversal} (CWE-22) endure maintainability
in $41.9\%$, $52.5\%$, $42.9\%$, $49.4\%$, and $41.7\%$, respectively. Therefore, 
we draw attention for the need of better documentation explaining how to use best practices to
produce secure and maintainable code; new and better programming languages; new
academic curricula in the fields of security and software quality; and, tools
such as BCH to help developers predict the effect of their patches.

As future work, this study may be extended in several directions: 
investigate which guidelines affect most the maintainability per
weakness; check if vulnerability patches are followed by refactoring
commits or not and how much time does it take to do it; 
investigate how patches impact maintainability for each
vulnerability and programming language; expand our methodology with other
software quality properties; validate these findings with closed or private
software; and, expand this analysis to other quality standards.

>>>>

This work presents an empirical study on the impact of $998$ security
patches on the maintainability of $262$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in previous work~\cite{Olivari:2018, cruz2019energyoriented}. Security patches 
significantly affect the software maintainability of open-source software, i.e., 
developers hinder maintainability while patching vulnerabilities. We found this evidence in
$41.28\%$ of the studied cases. Regarding regular commits, it seems that overall
maintainability stays the same ($42.69\%$). However, there is still a 
considerable amount of cases that harm maintainability ($27.05\%$).
Regarding security patches, we also observe that some guidelines are more
affected than others: \emph{Write Simple Units of Code} ($37.3\%$) and
\emph{Write Short Units of Code} ($37.6\%$) the most affected ones.

This study also shows evidence that there are security patterns that need more
attention than others. \emph{Improper Input Validation} (CWE-20), \emph{Information 
Exposure} (CWE-200), \emph{Improper Restriction of Operations within the Bounds of a 
Memory Buffer} (CWE-119), \emph{Missing Release of Memory after Effective Lifetime} 
(CWE-401) and \emph{Path Traversal} (CWE-22) endure maintainability
in $41.9\%$, $52.5\%$, $42.9\%$, $49.4\%$, and $41.7\%$, respectively. Therefore, 
we draw attention for the need of better documentation explaining 
how to use best practices to produce secure and maintainable code; 
new and better programming languages; new
academic curricula in the fields of security and software quality; and, 
tools such as Better Code Hub and efficient static analysis to help 
developers predict the effect of their patches on software maintainability.

As future work, this study may be extended in several directions: 
investigate which guidelines affect most the maintainability per
weakness; check if vulnerability patches are followed by refactoring
commits and how much time does it take to do it; 
expand our methodology with other software quality properties; 
validate these findings with closed/private
software; and, expand this analysis to other quality standards.
